### 2021-06-27 submission/manuscript.tex
Incorporate edits from coauthors

Antibiotics are a major risk factor for \emph{Clostridioides difficile}
infections (CDIs) because of their impact on the microbiota. However,
non-antibiotic medications such as the ubiquitous osmotic laxative
polyethylene glycol (PEG) 3350 also alter the microbiota. But whether
PEG impacts CDI susceptibility and clearance is unclear. To examine how
PEG impacts susceptibility, we treated C57Bl/6 mice with 5-day and 1-day
doses of 15\% PEG in the drinking water and then challenged the mice
with \emph{C. difficile} 630. We used clindamycin-treated mice as a
control because they consistently clear \emph{C. difficile} within 10
days post-challenge. PEG treatment alone was sufficient to render mice
susceptible and 5-day PEG-treated mice remained colonized for up to 30
days post-challenge. In contrast, 1-day PEG treated mice were
transiently colonized, clearing \emph{C. difficile} within 7 days
post-challenge. To examine how PEG treatment impacts clearance, we
administered a 1-day PEG treatment to clindamycin-treated, \emph{C.
difficile}-challenged mice. Administering PEG to mice after \emph{C.
difficile} challenge prolonged colonization up to 30 days
post-challenge. When we trained a random forest model with community
data from 5 days post-challenge, we were able to predict which mice
would exhibit prolonged colonization (AUROC = 0.90). Examining the
dynamics of these bacterial populations during the post-challenge period
revealed patterns in the relative abundances of \emph{Bacteroides},
\emph{Enterobacteriaceae}, \emph{Porphyromonadaceae},
\emph{Lachnospiraceae}, and \emph{Akkermansia} that were associated with
prolonged \emph{C. difficile} colonization in PEG-treated mice. Thus,
the osmotic laxative, PEG, rendered mice susceptible to \emph{C.
difficile} colonization and hindered clearance.

@@

Diarrheal samples from patients taking laxatives are typically rejected
for \emph{Clostridiodes difficile} testing. However, there are
similarities between the bacterial communities from people with diarrhea
or \emph{C. difficile} infections (CDI) including lower diversity
compared to communities from healthy patients. This observation led us
to hypothesize that diarrhea may be an indicator of \emph{C. difficile}
susceptibility. We explored how osmotic laxatives disrupt the
microbiota's colonization resistance to \emph{C. difficile} by
administering a laxative to mice either before or after \emph{C.
difficile} challenge. Our findings suggest that osmotic laxatives
disrupt colonization resistance to \emph{C. difficile}, and prevent
clearance among mice already colonized with \emph{C. difficile}.
Considering that most hospitals recommend not performing \emph{C.
difficile} testing on patients taking laxatives and laxatives are used
when administering fecal microbiota transplants via colonoscopy to
patients with recurrent CDIs, further studies are needed to evaluate if
laxatives impact microbiota colonization resistance in humans.

@@

Many of the non-antibiotic medications associated with CDIs are known to
modulate gastrointestinal motility leading to either increased or
decreased colonic transit time, which in turn also strongly impacts
microbiota composition and function (10, 11). Stool consistency often
serves as an approximation of intestinal motility (10). Our group has
shown that when \emph{C. difficile} negative samples from patients were
separated into two groups based on stool consistency, there were similar
microbiota features between samples from CDI patients and \emph{C.
difficile} negative patients with diarrhea compared to non-diarrheal
samples that were \emph{C. difficile} negative (12). The similar
community features between CDI patients and patients with diarrhea
included low alpha diversity and there were also only 6 bacterial taxa
had higher relative abundances in communities from CDI patients. These
results led to the hypothesis that bacterial communities from patients
experiencing diarrhea are susceptible to developing CDIs, regardless of
how they developed diarrhea.

@@

Beyond susceptibility, PEG is also relevant in the context of treating
recurrent CDIs via fecal microbiota transplant (FMT) where a healthy
microbiota is administered to the patient to restore colonization
resistance. For FMTs that are delivered via colonoscopy, patients
typically undergo bowel preparation by taking an osmotic laxative prior
to the procedure. Many of the FMT studies to date rationalize the use of
laxatives prior to the FMT (20--22) based on a 1996 case study with 2
pediatric patients where the authors suggested in the discussion that
the laxative may help flush \emph{C. difficile} spores and toxins from
the intestine (23).

@@

\textbf{5-day laxative treatment differentially disrupted the fecal
microbiota compared to clindamycin treatment.} Since osmotic laxatives
and clindamycin have previously been shown to disrupt the murine
microbiota (14--17), we hypothesized the different \emph{C. difficile}
colonization dynamics between mice treated with the osmotic laxative or
clindamycin were due to the two drugs having differential effects on the
microbiota. We profiled the stool microbiota over time by sequencing the
V4 region of the 16S rRNA gene to compare changes across treatment
groups. We found time (R\textsuperscript{2} = 0.29) and treatment group
(R\textsuperscript{2} = 0.21) explained half of the observed variation
between fecal communities with most of the remaining variation explained
by interactions between treatment group and other experimental variables
including time, cage, and sequencing preparation plate (PERMANOVA
combined R\textsuperscript{2} = 0.95, \emph{P} \textless{} 0.001, Fig.
2A, Data Set S1, sheet 1). None of the treatment groups recovered to
their baseline community structure either 10 or 30 days post-challenge
suggesting other community features besides recovery to baseline were
responsible for the prolonged \emph{C. difficile} colonization in
PEG-treated mice (Fig. 2B).

@@

Because \emph{C. difficile} was not immediately detectable in the stools
of the PEG-treated mice that were allowed to recover for 10 days prior
to challenge, we decided to examine if there were genera that changed
during the post-challenge period. We compared the communities from when
\emph{C. difficile} shifted from undetectable at 1 day post-challenge to
detectable in the stool samples with the density stabilizing around 8
days post-challenge (Fig. S1A). We found no genera with relative
abundances that were significantly different over the two time points
(Data Set S1, sheet 5). However, there was also wide variation between
individual mice regarding when \emph{C. difficile} became detectable
(Fig. S1A) as well as the relative abundances of bacterial genera in the
communities (Fig. S1B). For example, two mice had a high relative
abundance of \emph{Enterobacteriaceae} throughout the post-challenge
period. One mouse died on the sixth day post-challenge and in the other
\emph{C. difficile} was present at a high density from the 4th day
post-challenge onward (Fig. S1B). While we did not identify a clear
signal to explain the delayed appearance of \emph{C. difficile} in the
5-day PEG mice that were allowed to recover for 10 days prior to
challenge, the delay was striking and could reflect changes in microbial
activity or metabolites that were not examined in this study.

@@

Because there were differences in the mucosal microbiota including
detectable \emph{C. difficile} sequences in tissues from PEG-treated
mice relative to mice treated with clindamycin, we next examined the
severity of \emph{C. difficile} challenge by evaluating cecum and colon
histopathology (27). However, we found there was no difference in cecum
and colon scores between clindamycin and PEG-treated mice that were
challenged with \emph{C. difficile} at 4 days post-challenge (Fig. 3D),
the time point typically examined in \emph{C. difficile} 630 challenged
mice (28). We also looked at 6 days post-challenge because that was when
there was a large difference in \emph{C. difficile} density between PEG-
and clindamycin-treated mice (Fig. 1C). Although there was a slight
difference in the histopathology score of the colon between PEG and
clindamycin-treated mice, there was not a signifant difference in the
cecum and the overall score was relatively low (1.5 to 2.5 out of 12,
Fig. 3E). Therefore, although PEG treatment had a disruptive effect on
the mucosal microbiota, the impact of \emph{C. difficile} challenge on
the cecum and colon was similar between PEG and clindamycin treated
mice.

@@

While the disruptive effect of antibiotics on \emph{C. difficile}
colonization resistance is well established, the extent to which other
drugs such as laxatives disrupt colonization resistance was unclear. By
following mice treated with an osmotic laxative over time, we found that
a 5-day PEG treatment before challenge resulted in prolonged \emph{C.
difficile} colonization, while a 1-day PEG treatment resulted in
transient colonization without the use of antibiotics. The differences
in \emph{C. difficile} colonization dynamics between the 5- and 1-day
PEG treated mice were associated with differences in how much the
treatments disrupted the microbiota. Additionally, the intestinal
communities of 5-day PEG treated mice did not regain colonization
resistance after a 10-day recovery period. In contrast to the other
5-day PEG treatment groups, \emph{C. difficile} was not immediately
detectable in the stools of most of the mice in the 10-day recovery
group. We also examined the impact of PEG treatment after \emph{C.
difficile} challenge. In opposition to the hypothesis suggested by the
literature, we found that PEG treatment prolonged colonization relative
to mice that only recieved clindamycin treatment. We identified patterns
in the relative abundances of \emph{Bacteroides},
\emph{Enterobacteriaceae}, \emph{Akkermansia},
\emph{Porphyromonadaceae}, and \emph{Lachnospiraceae} that were
associated with prolonged \emph{C. difficile} colonization (Fig. 8).
Overall, our results demonstrated that osmotic laxative treatment alone
rendered mice susceptible to \emph{C. difficile} colonization and the
duration of colonization depended on the length of PEG treatment and
whether treatment was administered before or after challenge.

In addition to altering composition, laxative treatment may alter
microbiota-produced metabolites. A previous study demonstrated that a
5-day treatment of 10\% PEG depleted acetate and butyrate and increased
succinate compared to untreated mice (15). While we did not perform
metabolomic analysis, we did see bacteria known to produce benefical
metabolites were depleted in mice that cleared \emph{C. difficile}
compared to mice with prolonged colonization. For example,
\emph{Oscillibacter valericigenes} can produce the SCFA valerate (30),
and separate studies demonstrated valerate is depleted after clindamycin
treatment and inhibited \emph{C. difficile} growth \emph{in vitro} and
in C57BL/6 mice (31, 32). Similarly, \emph{Acetatifactor} can produce
acetate and butyrate (33), SCFAs that are decreased in mice with
prolonged \emph{C. difficile} infection after antibiotic treatment (34).
Thus protective bacteria and their metabolites could be depleted by
osmotive laxative treatment depending on the timing and duration of
treatment.

One possible explanation for the prolonged \emph{C. difficile}
colonization in 5-day PEG treated mice, might be due to the bacteria's
persistence in the mucosal compartment. In fact, it has been
hypothesized that \emph{C. difficile} biofilms may serve as reservoirs
for recurrent infections (35) and \emph{C. difficile} biofilms in the
mucus layer were recently identified in patients as aggregates with
\emph{Fusobacterium nucleatum} (36). There was an interesting pattern of
increased \emph{Enterobacteriaceae}, \emph{Bacteroides}, and \emph{C.
difficile} in both the stool and mucosal communities of PEG-treated mice
suggesting a potential synergy. \emph{Bacteroides} has the potential to
degrade mucus and the osmotic laxative may have allowed
\emph{Bacteroides} to colonize the mucosal niche by degrading mucin
glycans with glycosyl hydrolases that are absent in \emph{C. difficile}
(37). \emph{Bacteroides} persistence in the mucosal tissue might also
have helped \emph{Enterobacteriaceae} to colonize the region, as a
synergy between mucus-degrading \emph{B. fragilis} and \emph{E. coli}
has previously been described (38). A separate study demonstrated
\emph{C. difficile} was present in the outer mucus layer and associated
with \emph{Enterobacteriaceae} and \emph{Bacteroidaceae} using
fluorescent in situ hybridization (FISH) staining (39). However,
protective roles for \emph{Bacteroides} have also been demonstrated. For
example, \emph{B. fragilis} prevented CDI morbidity in a mouse model and
inhibited \emph{C. difficile} adherence \emph{in vitro} (40). In
coculture experiments \emph{B. longum} decreased \emph{C. difficile}
biofilm formation while \emph{B. thetaiotamicron} enhanced biofilm
formation (41) and \emph{B. dorei} reduced \emph{C. difficile} growth in
a 9-species community \emph{in vitro} (42). Therefore, whether
\emph{Bacteroides} is detrimental or beneficial in the context of
\emph{C. difficile} infection or colonization is still unclear, but the
niche and interactions with other bacteria may contribute.

\emph{Akkermansia} is also a mucin degrader with potentially beneficial
or detrimental roles depending on context in other diseases (43, 44). In
our study the relative abundance of \emph{Akkermansia} shifted over time
between groups of mice that either cleared \emph{C. difficile} or had
prolonged colonization. In the stool it was initially increased in mice
that cleared \emph{C. difficile}, but shifted after 5-days
post-challenge so that it was increased in mice that had prolonged
colonization. In the context of CDIs, some studies suggest a protective
role (45, 46), while others suggest a detrimental role because
\emph{Akkermansia} was positively correlated with \emph{C. difficile}
(47--50). Because the relative abundance of \emph{Akkermansia} was
dynamic in our study, it is unclear whether \emph{Akkermansia} helps
with clearance of \emph{C. difficile} or allows it to persist. A better
understanding how \emph{C. difficile} interacts with the mucosal
microbiota may lead to insights into CDIs, asymptomatic \emph{C.
difficile} carriage, and colonizatiion resistance.

>>>>

Antibiotics are a major risk factor for \emph{Clostridioides difficile}
infections (CDIs) because of their impact on the microbiota. However,
non-antibiotic medications such as the ubiquitous osmotic laxative
polyethylene glycol (PEG) 3350 also alter the microbiota. Clinicians
also hypothesize that PEG helps clear \emph{C. difficile}. But whether
PEG impacts CDI susceptibility and clearance is unclear. To examine how
PEG impacts susceptibility, we treated C57Bl/6 mice with 5-day and 1-day
doses of 15\% PEG in the drinking water and then challenged the mice
with \emph{C. difficile} 630. We used clindamycin-treated mice as a
control because they consistently clear \emph{C. difficile} within 10
days post-challenge. PEG treatment alone was sufficient to render mice
susceptible and 5-day PEG-treated mice remained colonized for up to 30
days post-challenge. In contrast, 1-day PEG treated mice were
transiently colonized, clearing \emph{C. difficile} within 7 days
post-challenge. To examine how PEG treatment impacts clearance, we
administered a 1-day PEG treatment to clindamycin-treated, \emph{C.
difficile}-challenged mice. Administering PEG to mice after \emph{C.
difficile} challenge prolonged colonization up to 30 days
post-challenge. When we trained a random forest model with community
data from 5 days post-challenge, we were able to predict which mice
would exhibit prolonged colonization (AUROC = 0.90). Examining the
dynamics of these bacterial populations during the post-challenge period
revealed patterns in the relative abundances of \emph{Bacteroides},
\emph{Enterobacteriaceae}, \emph{Porphyromonadaceae},
\emph{Lachnospiraceae}, and \emph{Akkermansia} that were associated with
prolonged \emph{C. difficile} colonization in PEG-treated mice. Thus,
the osmotic laxative, PEG, rendered mice susceptible to \emph{C.
difficile} colonization and hindered clearance.

@@

Diarrheal samples from patients taking laxatives are typically rejected
for \emph{Clostridiodes difficile} testing. However, there are
similarities between the bacterial communities from people with diarrhea
or \emph{C. difficile} infections (CDI) including lower diversity
compared to communities from healthy patients. This observation led us
to hypothesize that diarrhea may be an indicator of \emph{C. difficile}
susceptibility. We explored how osmotic laxatives disrupt the
microbiota's colonization resistance to \emph{C. difficile} by
administering a laxative to mice either before or after \emph{C.
difficile} challenge. Our findings suggest that osmotic laxatives
disrupt colonization resistance to \emph{C. difficile}, and prevent
clearance among mice already colonized with \emph{C. difficile}.
Considering that most hospitals recommend not performing \emph{C.
difficile} testing on patients taking laxatives and laxatives are
prescribed prior to administering fecal microbiota transplants via
colonoscopy to patients with recurrent CDIs, further studies are needed
to evaluate if laxatives impact microbiota colonization resistance in
humans.

@@

Many of the non-antibiotic medications associated with CDIs are known to
modulate gastrointestinal motility leading to either increased or
decreased colonic transit time, which in turn also strongly impacts
microbiota composition and function (10, 11). Stool consistency often
serves as an approximation of intestinal motility (10). Our group has
shown that when \emph{C. difficile} negative samples from patients were
separated into two groups based on stool consistency, there were similar
microbiota features between samples from CDI patients and \emph{C.
difficile} negative patients with diarrhea compared to non-diarrheal
samples that were \emph{C. difficile} negative (12). The similar
community features between CDI patients and patients with diarrhea
included low alpha diversity and only 6 bacterial taxa with higher
relative abundances in communities from CDI patients. These results led
to the hypothesis that bacterial communities from patients experiencing
diarrhea are susceptible to developing CDIs, regardless of how they
developed diarrhea.

@@

Beyond susceptibility, PEG is also relevant in the context of treating
recurrent CDIs via fecal microbiota transplant (FMT), where a healthy
microbiota is administered to the patient to restore colonization
resistance. For FMTs that are delivered via colonoscopy, patients
typically undergo bowel preparation by taking an osmotic laxative prior
to the procedure. Many of the FMT studies to date rationalize the use of
laxatives prior to the FMT (20--22) based on a 1996 case study with 2
pediatric patients where the authors suggested in the discussion that
the laxative may help flush \emph{C. difficile} spores and toxins from
the intestine (23).

@@

\textbf{5-day laxative treatment differentially disrupted the fecal
microbiota compared to clindamycin treatment.} Since osmotic laxatives
and clindamycin have previously been shown to disrupt the murine
microbiota (14--17), we hypothesized the different \emph{C. difficile}
colonization dynamics between mice treated with the osmotic laxative or
clindamycin were due to the two drugs having differential effects on the
microbiota. We profiled the stool microbiota over time by sequencing the
V4 region of the 16S rRNA gene to compare changes across treatment
groups. We found time (R\textsuperscript{2} = 0.29) and treatment group
(R\textsuperscript{2} = 0.21) explained half of the observed variation
between fecal communities with most of the remaining variation explained
by interactions between treatment group and other experimental variables
including time, cage, and sequencing preparation plate (PERMANOVA
combined R\textsuperscript{2} = 0.95, \emph{P} \textless{} 0.001, Fig.
2A, Data Set S1, sheet 1). None of the treatment groups recovered to
their baseline community structure either 10 or 30 days post-challenge,
suggesting other community features besides recovery to baseline were
responsible for the prolonged \emph{C. difficile} colonization in
PEG-treated mice (Fig. 2B).

@@

Because \emph{C. difficile} was not immediately detectable in the stools
of the PEG-treated mice that were allowed to recover for 10 days prior
to challenge, we decided to examine if there were genera that changed
during the post-challenge period. We compared the communities from when
\emph{C. difficile} shifted from undetectable at 1 day post-challenge to
detectable in the stool samples with the density stabilizing around 8
days post-challenge (Fig. S1A). We found no genera with relative
abundances that were significantly different over the two time points
(Data Set S1, sheet 5). However, there was also wide variation between
individual mice regarding when \emph{C. difficile} became detectable
(Fig. S1A) as well as the relative abundances of bacterial genera in the
communities (Fig. S1B). For example, two mice had a high relative
abundance of \emph{Enterobacteriaceae} throughout the post-challenge
period. One mouse died on the sixth day post-challenge and in the other,
\emph{C. difficile} was present at a high density from the 4th day
post-challenge onward (Fig. S1B). While we did not identify a clear
signal to explain the delayed appearance of \emph{C. difficile} in the
5-day PEG mice that were allowed to recover for 10 days prior to
challenge, the delay was striking and could reflect changes in microbial
activity or metabolites that were not examined in this study.

@@

Because there were differences in the mucosal microbiota, including
detectable \emph{C. difficile} sequences in tissues from PEG-treated
mice relative to mice treated with clindamycin, we next examined the
severity of \emph{C. difficile} challenge by evaluating cecum and colon
histopathology (27). However, we found there was no difference in cecum
and colon scores between clindamycin and PEG-treated mice that were
challenged with \emph{C. difficile} at 4 days post-challenge (Fig. 3D),
the time point typically examined in \emph{C. difficile} 630 challenged
mice (28). We also looked at 6 days post-challenge because that was when
there was a large difference in \emph{C. difficile} density between PEG-
and clindamycin-treated mice (Fig. 1C). Although there was a slight
difference in the histopathology score of the colon between PEG and
clindamycin-treated mice, there was not a signifant difference in the
cecum and the overall score was relatively low (1.5 to 2.5 out of 12,
Fig. 3E). Therefore, although PEG treatment had a disruptive effect on
the mucosal microbiota, the impact of \emph{C. difficile} challenge on
the cecum and colon was similar between PEG and clindamycin treated
mice.

@@

While the disruptive effect of antibiotics on \emph{C. difficile}
colonization resistance is well established, the extent to which other
drugs such as laxatives disrupt colonization resistance was unclear. By
following mice treated with an osmotic laxative over time, we found that
a 5-day PEG treatment before challenge resulted in prolonged \emph{C.
difficile} colonization, while a 1-day PEG treatment resulted in
transient colonization without the use of antibiotics. The differences
in \emph{C. difficile} colonization dynamics between the 5- and 1-day
PEG treated mice were associated with differences in the degree to which
treatments disrupted the microbiota. Additionally, the intestinal
communities of 5-day PEG treated mice did not regain colonization
resistance after a 10-day recovery period. In contrast to the other
5-day PEG treatment groups, \emph{C. difficile} was not immediately
detectable in the stools of most of the mice in the 10-day recovery
group. We also examined the impact of PEG treatment after \emph{C.
difficile} challenge. In opposition to the hypothesis suggested by the
literature, we found that PEG treatment prolonged colonization relative
to mice that only recieved clindamycin treatment. We identified patterns
in the relative abundances of \emph{Bacteroides},
\emph{Enterobacteriaceae}, \emph{Akkermansia},
\emph{Porphyromonadaceae}, and \emph{Lachnospiraceae} that were
associated with prolonged \emph{C. difficile} colonization (Fig. 8).
Overall, our results demonstrated that osmotic laxative treatment alone
rendered mice susceptible to \emph{C. difficile} colonization and the
duration of colonization depended on the length of PEG treatment and
whether treatment was administered before or after challenge.

In addition to altering composition, laxative treatment may alter
microbiota-produced metabolites. A previous study demonstrated that a
5-day treatment of 10\% PEG depleted acetate and butyrate and increased
succinate compared to untreated mice (15). While we did not perform
metabolomic analysis, we did see bacteria known to produce beneficial
metabolites were depleted in mice that cleared \emph{C. difficile}
compared to mice with prolonged colonization. For example,
\emph{Oscillibacter valericigenes} can produce the SCFA valerate (30),
and separate studies demonstrated valerate is depleted after clindamycin
treatment and inhibited \emph{C. difficile} growth \emph{in vitro} and
in C57BL/6 mice (31, 32). Similarly, \emph{Acetatifactor} can produce
acetate and butyrate (33), SCFAs that are decreased in mice with
prolonged \emph{C. difficile} infection after antibiotic treatment (34).
Thus protective bacteria and their metabolites could be depleted by
osmotive laxative treatment depending on the timing and duration of
treatment.

One possible explanation for the prolonged \emph{C. difficile}
colonization in 5-day PEG treated mice, might be due to the bacteria's
persistence in the mucosal compartment. In fact, it has been
hypothesized that \emph{C. difficile} biofilms may serve as reservoirs
for recurrent infections (35) and \emph{C. difficile} biofilms in the
mucus layer were recently identified in patients as aggregates with
\emph{Fusobacterium nucleatum} (36). There was an interesting pattern of
increased \emph{Enterobacteriaceae}, \emph{Bacteroides}, and \emph{C.
difficile} in both the stool and mucosal communities of PEG-treated mice
suggesting a potential synergy. \emph{Bacteroides} has the potential to
degrade mucus and the osmotic laxative may have allowed
\emph{Bacteroides} to colonize the mucosal niche by degrading mucin
glycans with glycosyl hydrolases that are absent in \emph{C. difficile}
(37). \emph{Bacteroides} persistence in the mucosal tissue might also
have helped \emph{Enterobacteriaceae} to colonize the region, as a
synergy between mucus-degrading \emph{B. fragilis} and \emph{E. coli}
has previously been described (38). A separate study demonstrated
\emph{C. difficile} was present in the outer mucus layer and associated
with \emph{Enterobacteriaceae} and \emph{Bacteroidaceae} using
fluorescent in situ hybridization (FISH) staining (39). However,
protective roles for \emph{Bacteroides} have also been demonstrated. For
example, \emph{B. fragilis} prevented CDI morbidity in a mouse model and
inhibited \emph{C. difficile} adherence \emph{in vitro} (40). In
coculture experiments, \emph{B. longum} decreased \emph{C. difficile}
biofilm formation while \emph{B. thetaiotamicron} enhanced biofilm
formation (41) and \emph{B. dorei} reduced \emph{C. difficile} growth in
a 9-species community \emph{in vitro} (42). Therefore, whether
\emph{Bacteroides} is detrimental or beneficial in the context of
\emph{C. difficile} infection or colonization is still unclear, but the
niche and interactions with other bacteria may contribute.

\emph{Akkermansia} is also a mucin degrader with potentially beneficial
or detrimental roles depending on context in other diseases (43, 44). In
our study, the relative abundance of \emph{Akkermansia} shifted over
time between groups of mice that either cleared \emph{C. difficile} or
had prolonged colonization. In the stool it was initially increased in
mice that cleared \emph{C. difficile}, but shifted after 5-days
post-challenge so that it was increased in mice that had prolonged
colonization. In the context of CDIs, some studies suggest a protective
role (45, 46), while others suggest a detrimental role because
\emph{Akkermansia} was positively correlated with \emph{C. difficile}
(47--50). Because the relative abundance of \emph{Akkermansia} was
dynamic in our study, it is unclear whether \emph{Akkermansia} helps
with clearance of \emph{C. difficile} or allows it to persist. A better
understanding how \emph{C. difficile} interacts with the mucosal
microbiota may lead to insights into CDIs, asymptomatic \emph{C.
difficile} carriage, and colonizatiion resistance.
